---
 title: "Geocode Steps"
author: "Hector Corrales"
date: "February 6, 2017"
output: html_document
---

1) Install necessary packages. DO NOT RUN THIS CHUNK IF YOU HAVE ALREADY RUN IT.
```{r}
#### Setup Libraries and Data. Note, you only need to install packages once. 
#### After that, you can delete those lines and keep only the ones that start with "library() ####

install.packages("devtools")
library(devtools)
devtools::install_github("dkahle/ggmap")

install.packages("plyr")
install.packages("maptools")
install.packages("dplyr")
install.packages("sp")
install.packages("rjson")
install.packages("tigris")
install.packages("RCurl")
install.packages("rgdal")
install.packages("httr")
install.packages("leaflet")
install.packages("reshape2")
install.packages("acs")
install.packages("mapview")
```

2) Load the Packages
```{r}

library(ggmap)
library(plyr)
library(maptools)
library(dplyr)
library(sp)
library(rjson)
library(tigris)
library(RCurl)
library(rgdal)
library(httr)
library(leaflet)
library(reshape2)
library(acs)
library(mapview)

```

3) Read in your data and GIS File
```{r}
dat <- read.csv("data/mydata.csv") # Read in your file with Addresses.

nys <- readShapePoly("data/tl_2016_36_tract", proj4string=CRS("+proj=longlat +datum=WGS84"))
syr <- nys[nys$COUNTYFP == "067", ]

```

4) Register your API Key and then Geocode your addresses using Google Maps API.
The following chunk of code will take some time to run, depending on how many addresses you wish to geocode. Be patient.
DO NOT RUN THIS CHUNK IF YOU HAVE RUN IT BEFORE.
```{r}

my.api.key <- "AIzaSyB47ttVaobUsKkFj2L41LFUtunVEYi7LD8"

register_google(key = 'AIzaSyB47ttVaobUsKkFj2L41LFUtunVEYi7LD8',
                account_type = "premium", day_limit = 150000)



mydata.coordinates <- geocode(     
  paste(dat$address,               
        dat$city,                  
        dat$state,                 
        sep = ", "))

# This following line saves the table with coordinates as an excel file. You must remove the hashtag in order to save after you've gone through the geocoding process.

#write.csv(mydata.coordinates, "data/geocoded.csv", row.names = FALSE) 


```

5) Read in the coordinates you saved above:
```{r}

mydata.coordinates <- read.csv("data/geocoded.csv") 


```

6) Add the coordinates to your file:
```{r}
mydata.withcoords <- cbind(dat, mydata.coordinates)

```

7) Check which entries could not be geocoded and save an excel file:
```{r}

no.na <- na.omit(mydata.withcoords)
not.found <- setdiff(mydata.withcoords, no.na )

write.csv(not.found, "data/Addresses not found.csv") # OPEN THIS FILE TO SEE THE ADDRESSES THAT WEREN'T FOUND.

rm(dat, mydata.coordinates, no.na, not.found)


```


8) Setup, Convert files to GIS and get Census Tract information.
```{r}


mydata.withcoords <- na.omit(mydata.withcoords)
to.sp <- data.frame(lon = mydata.withcoords$lon, lat = mydata.withcoords$lat)
mydata.sp <-  SpatialPoints(to.sp, 
                            proj4string = CRS("+proj=longlat
                                              +datum=WGS84" 
                                              ) )

mydata.tracts <- over(mydata.sp, syr, returnList = FALSE ) # This gets the Census Tract for the address.
mydata.withcoords$tract <- mydata.tracts$NAME

```


9) Product 1: Your original file, with coordinates and Census Tracts. Excel file will be saved in the data folder.

```{r}
write.csv(mydata.withcoords, "data/mygeocodeddata.csv", row.names = FALSE)
mygisdata <- read.csv("data/mygeocodeddata.csv")

```

10) Product 2: A map with dots. Each point represents a client.

```{r}

coordinates(mydata.withcoords) <- ~ lon + lat

proj4string(mydata.withcoords) <- CRS("+proj=longlat
                                              +datum=NAD83 
                                              +no_defs 
                                              +ellps=GRS80 
                                              +towgs84=0,0,0") 






dot.map <- mapview(mydata.withcoords,
            map.types = c("Esri.WorldStreetMap",
                          "CartoDB.Positron",
                          "OpenStreetMap",
                          "CartoDB.DarkMatter"),
            color = "steelblue",
            lwd = 2.5,
            layer.name = "All")

dot.map




                   

```

11) Product C: A table with how many clients in each Census Tract

```{r}

# Aggregate how many entries per Census Tract to draw the heatmap.

mydata.list <- over(syr, mydata.sp, returnList = TRUE )
count <- sapply( mydata.list, length )
myclient.tracts <- data.frame( GEOID10=syr$NAME, count=count ) # THIS IS FOR THE HEATMAP

# Append the client count to the Shapefile.
df_merged <- geo_join(syr, myclient.tracts, "NAME", "GEOID10") 

#### SHow product C: A file with the census tract and the (count or rate) of what you want to measure.

mytract.data <- as.data.frame(df_merged)
mytract.data <- select(mytract.data, NAME, count) # this is the dataframe to show for product C.

View(mytract.data)

```


12) Product D: A heatmap of where your clients are located:

```{r}

#### Show Product D: A heatmap. ####

popup <- paste0("Census Tract: ", df_merged$NAME, "<br>", 
                "Number of Clients here: ", df_merged$count, ".", "<br>",
                "THE INFO I WANT TO SEE")

pal <- colorNumeric(
  palette = "Blues",
  domain = df_merged$count # This is where you choose which values you want to visualize per Census Tract
)

my.choro <- leaflet() %>%
  addProviderTiles("Esri.WorldStreetMap") %>%
  setView(lng=-76.13, lat=43.03, zoom = 12) %>%
  addPolygons(data = df_merged, 
              fillColor = ~pal(count), 
              color = "gray80", # you need to use hex colors
              fillOpacity = 0.7, 
              weight = 1.5, 
              smoothFactor = 0.2,
              popup = popup) %>%
  
  addLegend(pal = pal, 
            values = df_merged$count, # Change your desired values here as well
            position = "bottomright", 
            title = "# of Clients",
            labFormat = labelFormat(suffix = "")) 

my.choro


```




??) Make a map with multiple layers: This chunk is in development. Ignore for now.

```{r}

# Keep map we did above

dot.map


# Make second layer: Client with a criminal record.
filt.layer <- filter(no.na, crecord == 1)
coordinates(filt.layer) <- ~ lon + lat
proj4string(filt.layer) <- "+init=epsg:4326"

crecord <- mapview(filt.layer,
            map.types = c("Esri.WorldStreetMap",
                          "CartoDB.Positron",
                          "OpenStreetMap",
                          "CartoDB.DarkMatter"),
            color = "steelblue",
            lwd = 2.5,
            layer.name = "W/ Criminal Record")


# The map will now have two layers. Click the layers button to turn or off.
dot.map + criminals

```
















